---
title: "RECHARGE GJAM style"
output: html_document
date: "2023-11-29"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(here)
```

Format RECHARGE data for GJAM analysis for kicks

# Bring in the raw data

Downloaded these annotations directly from CoralNet, where there were files through T14 (not sure if they're annotated).

```{r}
# Annotations as downloaded off of CoranNet on Oct 3, 2023
oct3anno <- read_csv(here("data/annotations_2023Oct3_CoralNet.csv")) %>% 
  filter(Time_point != 99) %>% 
  select(-Date)

# we'll also bring in a key for dealing with many of the categories on coralnet (this is all from Tom's awesome code)
key <- read_csv(here("data/annotation_key_feb2023.csv")) %>% 
  mutate(Label = Taxa)

# and a metadata file for identifying plot treatments
trt_key <- read_csv(here("data/TRT_key.csv"))
```

# Clean up the data

Need to make the CoralNet files all uniform

```{r}
oct3anno %>% 
  # standardize these names
  mutate(
    Herb_TRT = case_when(
      Herb_TRT == "1x1" ~ "1X1",
      Herb_TRT == "3x3" ~ "3X3",
      Herb_TRT == "open" ~ "Open",
      TRUE ~ Herb_TRT
    )) %>% 
  # add column for block plot and herb information
  mutate(Block_Plot = str_c(Block, Plot, sep = ""),
         Block_Plot_Herb_TRT = str_c(Block_Plot, Herb_TRT, sep = "_"),
         Count = 1) -> annots
```

Use the annotation key to get coarser taxa designations

```{r}
# then join the key with the annotation and treatement data
annots %>% 
  full_join(key, by = "Label") %>% 
  full_join(trt_key) -> recharge

```

Look to see what did not get labeled:

```{r}
recharge %>% filter(is.na(Taxa)) # something called Dic_ver (will need to ask about this if we move forward with this analysis)
```

# Summarize at the plot level

```{r}
recharge %>% 
  group_by(Time_point, Block, Plot, Block_Plot, Herb_TRT, Taxa2, Nutrient_Trt, Disturbance_Trt) %>% 
  # get count of each taxa per plot
  summarize(Count = sum(Count)) %>% 
  # pivot wider so each taxa gets a column
  pivot_wider(names_from = Taxa2, values_from = Count) %>% 
  # replace NA values in the matrix with 0
  replace(is.na(.), 0) %>% 
  # get rid of points that are out
  select(-OUT)  %>% 
  ungroup() %>% 
  select(-`NA`) -> plots
  
```

Get percent cover for each taxa group

```{r}
# first column we want
startcol <- grep("Acanthastrea_spp", colnames(plots))

# last column we want
endcol <- grep("Cladophoropsis", colnames(plots))

# calculate the total cover
sum_cover <- cbind(plots, Plot_sum = rowSums(plots[, startcol:endcol])) # looks about right--over 200 valid (not NA or out) points per plot

# and percent of hard cover
perc_cover <- (sum_cover[,startcol:endcol]/(sum_cover$Plot_sum-sum_cover$Sand)) * 100

# then combine columns again for the full dataset
percHard <- cbind(plots[1:(startcol-1)], perc_cover) 
```


# Calculate changes

```{r}
percHard %>% 
  # the columns we might want to use in the model
  select(Time_point, Block, Plot, Block_Plot, Herb_TRT, Nutrient_Trt, Disturbance_Trt, Acropora_spp, Montipora_spp, Pavona, Pocillopora_spp, Porites_spp) %>% 
  pivot_wider(names_from = Time_point, values_from = c(Acropora_spp, Montipora_spp, Pavona, Pocillopora_spp, Porites_spp), values_fill = 0) %>% 
  mutate(Change_MontipHard12 = Montipora_spp_12-Montipora_spp_0,
         Change_AcropHard12 = Acropora_spp_12-Acropora_spp_0,
         Change_PocHard12 = Pocillopora_spp_12-Pocillopora_spp_0,
         Change_PavonaHard12 = Pavona_12-Pavona_0,
         Change_PoritesHard12 = Porites_spp_12-Porites_spp_0
         

         ) -> coral_change
```

# Make ymatrix

```{r}
coral_change %>% 
  mutate(Rows = paste0(Block_Plot, Herb_TRT, Nutrient_Trt, Disturbance_Trt)) %>% 
  select(Rows, Change_MontipHard12, Change_AcropHard12, Change_PocHard12, Change_PavonaHard12, Change_PoritesHard12) -> ymatrix
```


# Make xmatrix

```{r}
coral_change %>% 
  mutate(Rows = paste0(Block_Plot, Herb_TRT, Nutrient_Trt, Disturbance_Trt)) %>% 
  select(Rows, Block, Plot, Block_Plot, Herb_TRT, Nutrient_Trt, Disturbance_Trt) -> xmatrix
```


# Save as csvs

```{r}
write_csv(xmatrix, here("generated_data/rechargexdata.csv"))
write_csv(ymatrix, here("generated_data/rechargeydata.csv"))

```


